---
title: "hw03-julie-song"
author: "Julie Song"
date: "3/13/2018"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, fig.path = 'images/')
```

###2. Pipelines and Redirection

####2.1) 
**Write a pipeline to obtain the unique team names, and redirect the output to a text file team-names.txt. Use head to display the first five lines of the created file.**
```{bash}
cut -d "," -f 2 nba2017-roster.csv | tail +2 | sort |uniq > team-names.txt

head -5 team-names.txt
```

####2.2) 
**Write a pipeline to obtain the unique positions, and redirect the output to a text file position-names.txt. Use head to display the first five lines of the created file.**
```{bash}
cut -d "," -f 3 nba2017-roster.csv | tail +2 | sort | uniq > position-names.txt

head -5 position-names.txt
```

####2.3)
**Write a pipeline to obtain the counts of the different experience values, displayed in descending order. Redirect the output to a text file experience-counts.txt. Use head to display the first five lines of the created file. The first column corresponds to count, the second column corresponds to experience.**

```{bash}
cut -d "," -f 7 nba2017-roster.csv | tail +2 | sort | uniq -c | sort -nr > experience-counts.txt
head -5 experience-counts.txt
```

####2.4)
**Use output redirection commands to create a CSV file LAC.csv containing data for the LAC team (Los Angeles Clippers). Your CSV file should include column names. Use cat to display the content of the created file**
```{bash}
grep 'player' nba2017-roster.csv > LAC.csv
grep 'LAC' nba2017-roster.csv >> LAC.csv
cat LAC.csv
```

####2.5)
**Write a pipeline to display the age frequencies of LAL players. The first column corresponds to count, the second column corresponds to age.**
```{bash}
grep 'LAL' nba2017-roster.csv | cut -d "," -f 6 | sort | uniq -c
```

####2.6)
**Write a pipeline to find the number of players in CLE (Cleveland) team; the output should be just the number of players.**

```{bash}
grep 'CLE' nba2017-roster.csv | wc -l

```

####2.7)
**Write pipelines to create a CSV file gsw-height-weight.csv that contains the player, height and weight of GSW players. Your CSV file should include column names. Use cat to display the file contents**
```{bash}
grep 'player' nba2017-roster.csv | cut -d "," -f 1,4,5 > gsw-height-weight.csv
grep 'GSW' nba2017-roster.csv | cut -d "," -f 1,4,5 >> gsw-height-weight.csv 
cat gsw-height-weight.csv

```

####2.8) 
**Write pipelines to create a file top10-salaries.csv containing the top10 player salaries,arranged by salary from largest to smallest. Your CSV file should include column names. Use cat to display the file contents**

```{bash}
grep 'player' nba2017-roster.csv | cut -d "," -f 1,8 > top10-salaries.csv
sort -t "," -n -k8 -r nba2017-roster.csv | cut -d "," -f 1,8 | head -10 >> top10-salaries.csv

cat top10-salaries.csv
```

###3. Rmd File

```{r}
source('code/binomial-functions.R')
```

**Assume that the “successful” event is getting a “six” when rolling a die. Consider rolling a fair die 10 times. Use bin_probability() to find the probability of getting exactly 3 sixes.**

```{r}
bin_probability(10, 3, 1/6)
```

**Use bin_distribution() to obtain the distribution of the number of “sixes” when rolling a loaded die 10 times, in which the number “six” has probability of 0.25. Make a plot of this distribution.**

```{r}
plot(bin_distribution(10, 0.25), main = 'Scatterplot of Probability Distribution on a Biased Die (25% Sixes)', xlab = 'Number of Sixes', ylab = 'Probability of Getting Sixes')
```

**Use bin_probability(), and a for loop, to obtain the probability of getting more than 3 heads in 5 tosses with a biased coin of 35% chance of heads.**

```{r}
term <- 0
for(i in 4:5) {
  term <- term + bin_probability(5, i, 0.35)
}
term

```

**Use bin_distribution() to obtain the probability distribution of the number of heads when tossing a loaded coin 15 times, with 35% chance of heads. Make a plot of this distribution.**
```{r}
plot(bin_distribution(15, 0.35), main = 'Scatterplot of Probability Distribution on a Biased Coin (35% Heads)', xlab = 'Number of Heads', ylab = 'Probability of Getting Head(s)')

```


