---
title: "Getting Started with dieroller"
author: "Julie Song"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to dieroller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(dieroller)
```

##Die Rolling
The package "dieroller" is a minimal implementation for simulating rolling a six-sided standard die a given number of times, and to visualize the relative frequencies in such a series of rolls.

##Creating a die
The first step is to create a "coin" object with the function die():

```{r}
mydie <- die()
mydie

```

By default, die() creates a fair die with standard sides from 1 through 6, and with probability values for each side of 1/6.

You can change the defatuls with arguments sides and prob:

```{r}
loaded <- die(sides = c('i', 'ii', 'iii', 'iv', 'v', 'vi'), prob = c(0.1, 0.2, 0.1, 0.2, 0.1, 0.3))
loaded
```

##Rolling a die
Once you have defined a "die" object, you can use the function roll() to toss a die a given number of times. The output will be an object of class "roll", which will contain the vector of rolls:

```{r}
roll100 <- roll(mydie, times = 100)
roll100
```

The summary() function gives you a nice table of frequencies:

```{r}
summary(roll100)
```

##Plot of relative frequencies
You can use the function plot() to create a graohic with the relative frequencies of a series of coin tosses:

```{r}
plot(roll100)
```
